---
title: "Live from Tender Electronic Daily Database"
author: "Pierre-Henri Morand"
date: 'date: `r Sys.Date()`'
output:
  html_document:
    toc: yes
    toc_float: yes
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, warning = TRUE,  message = FALSE, cache.lazy = FALSE)

#knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)

```


# Presentation of the database

## Main characteristics

- The data comes from public procurement standard forms, which are filled in by
contracting bodies and sent as notices for publication to TED. Most of the variables in the data directly correspond to the fields in the forms.

- TED - Contract award notices 2010-2019. It is a merged file for the analysis across the years.

- TED data has multiple levels. Contract awards (CAs), which is the fifth section of a CAN standard form. Each CAN has one or more CAs. ID_award_ca is an ID of a CA.

- These levels serve different purposes. To give an example, a CAN generally informs on the result of the procurement (“The ministry has bought furniture”); and the three individual CAs, nested within the CAN, inform about the specific detail (“The ministry has bought a chair from company A”; “The ministry has bought a table from company B”; “The ministry has bought a cupboard from company B”). 

## Downloading the data

The latest database can be downloaded directly [here](https://data.europa.eu/euodp/repository/ec/dg-grow/mapps/TED%202020/Export_OpenDataCAN_from2010to2019.csv)


```{r data dwonloading, echo=FALSE, message=TRUE, warning=TRUE, cache=TRUE}
#importing useful library

library(tidyverse)
library(banR)
library(jsonlite)
library(magrittr)
library(data.tree)
library(geosphere)
library(sf)
library(cartography)
library(stringr)

#creating the data repository
data_dir <- 'data'
if (!dir.exists(data_dir)) {
  dir.create(data_dir)
}

# downloading the DECP JSON file from the remote data.europa.eu site.
TED_url <- "https://data.europa.eu/euodp/repository/ec/dg-grow/mapps/TED%202020/Export_OpenDataCAN_from2010to2019.csv"
TED_destfile <-  str_c(data_dir, "/Export_OpenDataCAN_from2010to2019.csv")
if (!file.exists(TED_destfile)) {
  download.file(TED_url, TED_destfile)
}

# creating the dataframe
TED<- read.csv(TED_destfile)
```

# Some data to work up the appetite!

The database represents `r length(TED$ID_NOTICE_CAN)` contrats awarded between 2010 and 2019. Here is the number of contracts by buyers country and by year :

```{r echo=FALSE,  cache=TRUE}
library(DT)
library(tidyverse)

#datatable(table(TED$ISO_COUNTRY_CODE, TED$YEAR), colnames = c("Country code", "Year", "Nb of Contracts"))
table_1<-as.data.frame(table(TED$ISO_COUNTRY_CODE, TED$YEAR))
table_1 %>%
  datatable(extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))), colnames = c("Country code", "Year", "Nb of Contracts"))
```

France (as well as Poland) accounts for a large share of the data available under TED.

```{r echo=FALSE,  cache=TRUE}
table_2<-table(TED[which(TED$ISO_COUNTRY_CODE=="FR" |TED$ISO_COUNTRY_CODE=="DE"|TED$ISO_COUNTRY_CODE=="IT"|TED$ISO_COUNTRY_CODE=="UK"|TED$ISO_COUNTRY_CODE=="PL"),]$YEAR, TED[which(TED$ISO_COUNTRY_CODE=="FR" |TED$ISO_COUNTRY_CODE=="DE"|TED$ISO_COUNTRY_CODE=="IT"|TED$ISO_COUNTRY_CODE=="UK"|TED$ISO_COUNTRY_CODE=="PL"),]$ISO_COUNTRY_CODE)
barplot(table_2, legend.text = TRUE, main="cumulative nb of contracts", ylab="Nb", xlab="Country")
```
  

The distribution of contracts by main categories gives us, for the entire database:

```{r, echo=FALSE, cache=TRUE}
library(RColorBrewer)
count<-table(TED$TYPE_OF_CONTRACT)
color <- brewer.pal(length(count), "Set2")
pie(table(TED$TYPE_OF_CONTRACT), labels = table(TED$TYPE_OF_CONTRACT), col = color, density = 50, angle = 45)
pie_labels <- paste0(c("Works: ", "Furnitures: ", "Services: "),round(100 * count/sum(count), 2), "%")
pie(count, labels = pie_labels,col = color, density = 50, angle = 45)
```

Similarly, the type of purchasers corresponds to:`
```{r, echo=FALSE, cache=TRUE}
library(DT)
library(tidyverse)

#datatable(table(TED$ISO_COUNTRY_CODE, TED$YEAR), colnames = c("Country code", "Year", "Nb of Contracts"))
table_2<-as.data.frame(table(TED$CAE_TYPE))
table_2 %>%
  datatable(extensions = 'Buttons',
            options = list(dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))), colnames = c("CAE type",  "Nb of Contracts"))
```
with:
  
  |CAE Code |CAE Type|
|:--------:|:----------|
 | 1 |  Ministry or any other national or federal authority, including their regional of local subdivisions|
|3 |Regional or local authority|
|4 |Water, energy, transport and telecommunications sectors|
|5 |European Union institution/agency|
|5A |other international organisation|
|6 |Body governed by public law|
|8 |Other|
|N |National or federal Agency / Office|
|R |Regional or local Agency / Office|
|Z |Not specified|
